---
title: "Patching the CWB: Departure"
author: "Andreas Blaette"
date: "7/9/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(git2r)
library(fs)
library(magrittr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(root.dir = path.expand("~/Lab/github/RcppCWB"))
```


```{r}
repodir <- path.expand("~/Lab/github/RcppCWB")
```

The aim is to reconstruct all changes to the CWB code so that we know which patches are
necessary for adapting a fresh copy of the CWB such that it conforms R package requirements.

The approach taken here is
  (a) to create a branch with the original CWB code in the ./src/cwb directory
  (b) to apply patches to this branch
  (c) to check differences between the 'r1069'-branch (with patches) and the 'dev'-branch

As a result, we should have a precise understanding which changes led from CWB 3.4.14 (r1069) 
to the modified CWB code we have on the current 'dev'-branch. We should also have a reproducible 
workflow that can also be applied to the copy of a recent version of the CWB. This is the
basis for integrating upstream CWB code into RcppCWB.

The first step (a) is to create a fresh copy of CWB 3.4.14 (r1069) on a branch
named after the revision. The 'dev' branch is the point of departure and we refrain from 
removing the src/cwb directory: This gives us a much better understanding which files 
have been added during the development of RcppCWB and which ones have been deleted.


```{r}
# A copy of the SVN repo of the CWB is assumed to be present here ...
cwb_dir_svn <- path.expand("~/Lab/tmp/cwb/trunk") # r1069
```
  

### Create new branch r1069


```{r}
git2r::checkout(repodir, branch = "dev")
if ("r1069" %in% names(branches(repodir))){
  git2r::branch_delete(branch = branches(repodir)[["r1069"]])
}
git2r::branch_create(commit = last_commit(repo = repodir), name = "r1069")
git2r::checkout(repodir, branch = "r1069")
```


### Copy files from SVN copy of CWB to RcppCWB repo

```{r}
cwb_dir_repo <- path(repodir, "src", "cwb")
cwb_files_svn <- list.files(cwb_dir_svn, recursive = TRUE, full.names = TRUE)
file.copy(
  from = cwb_files_svn,
  to = gsub(paste("^", cwb_dir_svn, sep = ""), cwb_dir_repo, cwb_files_svn),
  overwrite = TRUE
)
```  


### Remove untracked files (files that have been removed)

Use git to detect files that were deleted from the original CWB code. These files have been copied but 
are untracked.

```{r root.dir = path.expand("~/Lab/github/RcppCWB"), eval = FALSE}
removed_files <- status(repo = repodir, staged = FALSE, unstaged = FALSE, untracked = TRUE) %>%
  unlist() %>%
  unname() %>%
  grep("^src/cwb", ., value = TRUE) %>%
  path(repodir, .)
file.remove(removed_files)
removed_files
```

### Remove files that have been added (and that need to be added explicitly)

To restore the state of RcppCWB development, these files need to be re-added
or generated later on

```{r}
cwb_files_svn_truncated <- gsub(cwb_dir_svn, "", cwb_files_svn)
cwb_files_repo <- list.files(cwb_dir_repo, recursive = TRUE, full.names = TRUE)
cwb_files_repo_truncated <- gsub(cwb_dir_repo, "", cwb_files_repo)
added_files <-  cwb_files_repo_truncated[!cwb_files_repo_truncated %in% cwb_files_svn_truncated]
added_files_full_path <- path(cwb_dir_repo, added_files)
added_files_existing <- added_files_full_path[file.exists(added_files_full_path)]
rm_file(repo = repodir, path = added_files_existing)
```
  
  
### git add/commit so that we can perform diffs

```{r}
git2r::add(repo = repodir, path = unname(unlist(git2r::status(repo = repodir))))
git2r::commit(repodir, message = "CWB r1069 restored")
```


### Inspect diffs
  
The code of CWB 3.4.14 has been used first on June 5 2018 and the commit message 
was 'first try with v3.4.14'. Get the commit  ID

```{r}
commit_v3_4_14 <- "8f92126f525909347fd532d975afcf105b5ec8c0"
v3_4_14_commit <- commits(repo = repodir)[[grep("first try with v3.4.14", sapply(commits(repo = repodir), `[[`, "message"))]]
```


#### cl directory
  
```{r, root.dir = "/Users/andreasblaette/Lab/github/RcppCWB/src/cwb/cl"}
setwd("~/Lab/github/RcppCWB/src/cwb/cl")
git_cmd <- sprintf(
  "git diff %s..%s . %s",
  git2r::last_commit(repo = repodir)[["sha"]], v3_4_14_commit[["sha"]],
  "':!*endian.h'"
)
cat(paste(system(git_cmd, intern = TRUE), collapse = "\n"))
```
  Finding: Modifications of the Makefile, nothing else

#### cqp directory

```{r}  
setwd("~/Lab/github/RcppCWB/src/cwb/cqp")
git_cmd <- sprintf(
  "git diff %s..%s . %s",
  git2r::last_commit()[["sha"]], v3_4_14_commit[["sha"]],
  "':!*_options.h' ':!*parser.tab.c' ':!*parser.tab.h'"
)
cat(paste(system(git_cmd, intern = TRUE), collapse ="\n"))
```
  
Finding: The Makefile has beend modified, and functions hash_string() find_prime() is_prime() are commented out in hash.c


#### CQi directory

```{r}
setwd("~/Lab/github/RcppCWB/src/cwb/CQi")
cat(paste(system(git_cmd, intern = TRUE), collapse ="\n"))
```

Finding: Identity


#### utils directory
  
```{r}
setwd("~/Lab/github/RcppCWB/src/cwb/utils")
git_cmd <- sprintf("git diff %s..%s ./*.c", git2r::last_commit()[["sha"]], v3_4_14_commit[["sha"]])
cat(paste(system(git_cmd, intern = TRUE), collapse ="\n"))

git_cmd <- sprintf("git diff %s..%s ./*.h", git2r::last_commit()[["sha"]], v3_4_14_commit[["sha"]])
cat(paste(system(git_cmd, intern = TRUE), collapse ="\n"))

```

Finding: '#include <netinet/in.h>' added to fileutils.h

### Cleaning up

```{r}
checkout(repo = repodir, branch = "dev")
```
