---
title: "Patching the CWB: Diff report"
author: "Andreas Blaette"
date: "7/9/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

## About this document

We assume that the branch 'r1069' has been created with a clean CWB v3.4.14 and that patches have been applied using the patch.R script. The purpose of this document is to identify the remaining differences to the latest commit of the dev branch of RcppCWB.


## Apply patches

```{r}
library(git2r)
library(fs)
repodir <- path.expand("~/Lab/github/RcppCWB")

checkout(repodir, branch = "dev")
last_devbranch_commit <- last_commit(repo = repodir)

patchfile <- path(repodir, "prep", "patch.R")
tmp_patchfile <- path(tempdir(), "path.R")
file_copy(path = patchfile, new_path = tmp_patchfile, overwrite = TRUE)
```

```{r}
checkout(repodir, branch = "r1069")
source(tmp_patchfile)
```

```{r}
add(repo = repodir, path = "src/cwb/*")
commit(repo = repodir, message = "scripted cwb patch")
last_patchbranch_commit <- last_commit(repo = repodir)
```


## Inspect diffs

#### cl directory
  
```{r}
setwd("~/Lab/github/RcppCWB/src/cwb/cl")
git_cmd <- sprintf(
  "git diff %s..%s . %s",
  last_patchbranch_commit[["sha"]], last_devbranch_commit[["sha"]],
  "':!*endian.h'"
)
result <- system(git_cmd, intern = TRUE)
cat(paste(result, collapse = "\n"))
```
  Finding: Modifications of the Makefile, nothing else

#### cqp directory

```{r}  
setwd("~/Lab/github/RcppCWB/src/cwb/cqp")
git_cmd <- sprintf(
  "git diff %s..%s . %s",
  last_patchbranch_commit[["sha"]], last_devbranch_commit[["sha"]],
  "':!*_options.h' ':!*parser.tab.c' ':!*parser.tab.h'"
)
result <- system(git_cmd, intern = TRUE)
cat(paste(result, collapse = "\n"))
```


#### CQi directory

```{r}
setwd("~/Lab/github/RcppCWB/src/cwb/CQi")
git_cmd <- sprintf(
  "git diff %s..%s .",
  last_patchbranch_commit[["sha"]], last_devbranch_commit[["sha"]]
)
result <- system(git_cmd, intern = TRUE)
cat(paste(result, collapse = "\n"))
```


#### utils directory
  
```{r}
setwd("~/Lab/github/RcppCWB/src/cwb/utils")
git_cmd <- sprintf(
  "git diff %s..%s .",
  last_patchbranch_commit[["sha"]], last_devbranch_commit[["sha"]]
)
result <- system(git_cmd, intern = TRUE)
cat(paste(result, collapse = "\n"))
```
