---
title: "Untitled"
author: "Andreas Blaette"
date: "2 2 2018"
output: html_document
---

## cwb-encode

# word.corpus

The corpus file is a sequence of the token ids. The length of the corpus file will be the number of tokens in a corpus.

```{r}
corpusFile <- "~/Lab/tmp/encoding/indexed_corpora/word.corpus"
corpusSize <- file.info(corpusFile)$size
corpus <- readBin(con = corpusFile, what = integer(), size = 4L, n = lexiconSize, endian = "big")
```


# word.lexicon

The file 'word.lexicon' includes the unique words.

```{r}
lexiconFile <- "~/Lab/tmp/encoding/indexed_corpora/word.lexicon"
lexiconSize <- file.info(lexiconFile)$size
lexicon <- readBin(con = lexiconFile, what = character(), n = lexiconSize)
```

# word.lexicon.idx

The idx file offers the positions/beginnings of a token in the lexicon file.

```{r}
idxFile <- "~/Lab/tmp/encoding/indexed_corpora/word.lexicon.idx"
idxSize <- file.info(idxFile)$size
idx <- readBin(con = idxFile, what = integer(), size = 4L, n = idxSize, endian = "big")
```

To see that ...

```{r}
i <- strsplit(paste(lexicon, collapse = "|"), split = "")[[1]]
names(i) <- 0:(length(i) - 1)
```


## cwb-makeall

```{sh}
cwb-makeall -r ~/Lab/tmp/encoding/registry -V FOO
```

# word.corpus.cnt

The cnt file is a list of integer values, the frequency of the tokens.

```{r}
cntFile <- "~/Lab/tmp/encoding/indexed_corpora/word.corpus.cnt"
cntSize <- file.info(cntFile)$size
cnt <- readBin(con = cntFile, what = integer(), size = 4L, n = idxSize, endian = "big")
```

# word.corpus.rdx

This points to the position in the rev file where the corpus positions for a token id begin.

```{r}
rdxFile <- "~/Lab/tmp/encoding/indexed_corpora/word.corpus.rdx"
rdxSize <- file.info(rdxFile)$size
rdx <- readBin(con = rdxFile, what = integer(), size = 4L, n = rexSize, endian = "big")
```

# word.corpus.rev

As long as the number of tokens. Indicates the corpus positions of a token.

```{r}
revFile <- "~/Lab/tmp/encoding/indexed_corpora/word.corpus.rev"
revSize <- file.info(revFile)$size
rev <- readBin(con = revFile, what = integer(), size = 4L, n = revSize, endian = "big")
```

# word.lexicon.srt

```{r}
srtFile <- "~/Lab/tmp/encoding/indexed_corpora/word.lexicon.srt"
srtSize <- file.info(srtFile)$size
srt <- readBin(con = srtFile, what = integer(), size = 4L, n = srtSize, endian = "big")
```


## cwb-huffcode

```{sh}
cwb-huffcode -r ~/Lab/tmp/encoding/registry -A FOO
```

### word.hcd
### word.huf
### word.huf.syn

## cwb-compress-rdx

```{sh}
cwb-compress-rdx -r ~/Lab/tmp/encoding/registry -A FOO
```

### word.crc (compressed index, replaces word.corpus.rev)

### word.crx (compressed index offsets, replaces word.crx)

